### 2021-12-27 dedication/dedication.tex
finish dedication

\cleardoublepage

\begin{dedication}
  I hereby declare that the content of this thesis is the product of my own research work. The main parts of this thesis have been published in the following list of my publications. I made the major contribution in design, implementation and experiments for these work under the guidance of Prof. Anil Anthony Bharath and some others which have been acknowledged appropriately in text. The information, codes, and former ideas guided from other sources have been cited in a list of references is in the bibliography.
\end{dedication}
% List the publications used in the thesis
\textbf{Publication List}:
\begin{enumerate}
	\item \underline{Tianhong Dai}, Magda Dubois, Kai Arulkumaran, Jonathan Campbell, Cher Bass, Benjamin Billot, Fatmatulzehra Uslu, Vincenzo de Paola, Claudia Clopath, and Anil Anthony Bharath, ``Deep Reinforcement Learning for Subpixel Neural Tracking", International Conference on Medical Imaging with Deep Learning, 2019 (Spotlight).
	\item \underline{Tianhong Dai}, Hengyan Liu, Anil Anthony Bharath, ``Episodic Self-Imitation Learning with Hindsight", Electronics, Special Issue on Deep Reinforcement Learning: Methods and Applications, 2020.
	\item \underline{Tianhong Dai}, Hengyan Liu, Kai Arulkumaran, Guangyu Ren, and Anil Anthony Bharath, ``Diversity‐based Trajectory and Goal Selection with Hindsight Experience Replay", Pacific Rim International Conference on Artificial Intelligence, 2021.
	\item \underline{Tianhong Dai}, Yali Du, Meng Fang, and Anil Anthony Bharath, ``Diversity-Augmented Intrinsic Motivation for Deep Reinforcement Learning", Neurocomputing, 2021.
	\item \underline{Tianhong Dai}, Kai Arulkumaran, Tamara Gerbert, Samyakh Tukra, Feryal Behbahani, and Anil Anthony Bharath, ``Analysing Deep Reinforcement Learning Agents Trained with Domain Randomisation", Neurocomputing, 2022.
\end{enumerate}

% List extra publications
In addition to the above papers, these following publications that I authored/co-authored during my PhD are out of the main scope of this thesis, and are briefly mentioned in the background or future directions.
\begin{enumerate}
	\item Shafa Balaram, Kai Arulkumaran, \underline{Tianhong Dai}, and Anil Anthony Bharath, ``A Maximum Entropy Deep Reinforcement Learning Neural Tracker", International Workshop on Machine Learning in Medical Imaging, 2019.
	\item Yali Du, Lei Han, Meng Fang, \underline{Tianhong Dai}, Ji Liu, and Dacheng Tao, ``LIIR: Learning Individual Intrinsic Reward in Multi-Agent Reinforcement Learning", Advances in Neural Information Processing Systems, 2019.
	\item \underline{Tianhong Dai}, Wei Li, Xilei Cao, Jianzhuang Liu, Xu Jia, Ales Leonardis, Youliang Yan, and Shanxin Yuan, ``Wavelet-Based Network For High Dynamic Range Imaging", Submitted to IEEE Transactions on Imaging Processing, 2021.
\end{enumerate}