### 2022-04-23 acknowledgements/acknowledgements.tex

\cleardoublepage

\addcontentsline{toc}{chapter}{Acknowledgements}

\begin{acknowledgements}

I would like to express my deep gratitude to:

\begin{itemize}
 \item Anil Anthony Bharath: Anil is my supervisor and brings me into an academic research career. During my Ph.D. study, he provided me with plentiful guidance, help and trust in my research. In addition, he has also provided me with valuable suggestions on my career planning. Without him, I would not have been able to achieve the current results.
 \vspace*{3mm}
 \item Kai Arulkumaran: Kai is my closest colleague/collaborator in the BICI group. He has helped me a lot in my research. In addition, he also gives me many suggestions in daily life. He is also my ``leader" in the field of reinforcement learning.
 \vspace*{3mm}
 \item BICI Group Members: I would like to express my gratitude for their daily support.
 \vspace*{3mm}
 \item My parents: From the day I was born, my parents have provided me with unfailing love and support. Without them, I wouldn't be here today.
 \vspace*{3mm}
 \item My wife: From the day we meet, she continually inspires me to chase my dreams and gives me much love and care. During our two years together in the UK, she has been there to guide me through difficult times, to help me with my worries and to allow me to focus on my work. She makes my life so full of hope and joy that no words can express my gratitude and love to her.
 % 从我们认识的那天起，她就给予我充足的关怀以及爱。我们一起在英国的这两年间，她每天为我准备美味的食物以及提供舒适的生活环境，让我能够专注于我自己的工作。也是她让我的生活充满了希望与快乐，再多的文字也表达不出我对她的感激与爱。
\end{itemize}

\end{acknowledgements}