### 2022-04-23 chapters/chapter3.tex

\chapter{Episodic Self-Imitation Learning with Hindsight}
\label{ch:esil}
\section{Introduction}
% \td{In this chapter, we propose episodic self-imitation learning (ESIL), which aims to address }
%------------------------------ From Main Paper ----------------------------
Reinforcement learning (RL) has been shown to be very effective in training agents within gaming environments~\cite{mnih2015human,silver2016mastering}, particularly when combined with deep neural networks~\cite{lecun2015deep,silver2016mastering,liu2017survey}. In most tasks settings that are solved by RL algorithms, reward shaping is an essential requirement for guiding the learning of the agent. However, reward shaping often requires significant quantities of domain knowledge that {are} highly task-specific~\cite{ng1999policy} and, even with careful design, can lead to undesired policies. Moreover, for complex robot manipulation tasks, the manual design of reward shaping functions to guide the learning agent becomes intractable~\cite{arulkumaran2017deep,florensa2017reverse} if even minor variations are introduced in the task. For such settings, the application of deep reinforcement learning requires algorithms that can learn from unshaped, and usually sparse, reward signals. The complicated dynamics of robot manipulation exacerbates the difficulty posed by sparse rewards, especially for on-policy RL algorithms. For example, achieving goals that require the successful execution of multiple steps over a long horizon involves high-dimensional control that must also be generalised to work across variations in the environment for each step. These aspects of robot control result in a situation where a naive RL agent so rarely receives a reward at the start of training that it cannot learn at all. A common solution in the \td{DRL} community is collecting a sufficient number of expert demonstrations, then use imitation learning to train the agent. However, in some scenarios, demonstrations are expensive to collect and the performance of a trained agent is restricted by their quantity. One solution is to use the valuable past experiences of the agent to improve training, which is particularly useful in sparse reward environments.

There are two kinds of approaches to alleviate the problems associated with sparse rewards: imitation learning and hindsight experience replay (HER)~\cite{andrychowicz2017hindsight}. First, the standard approach of imitation learning is to use supervised learning algorithms and minimise a surrogate loss with respect to an oracle. The most common form is learning from demonstrations~\cite{hester2018deep,gao2018reinforcement}. Similar techniques are applied to robot manipulation tasks~\cite{rajeswaran2017learning,vevcerik2017leveraging,nair2018overcoming,james2018task}. When the demonstrations are not attainable, self-imitation learning (SIL)~\cite{oh2018self}, which uses past good experiences (episodes in which the goal is achieved), can~be used to enhance the exploration or speed up the training of the agent. {Self-imitation learning} works well in discrete control environments, such as Atari games. While being able to learn policies for continuous control tasks with dense or delayed rewards~\cite{oh2018self}, the present experiments suggest that SIL struggles when the rewards are sparse. Recently, {hindsight experience replay} has been proposed to solve such goal-oriented, sparse reward problems. The main idea of HER is that during replay, the selected transitions are sampled from state--action pairs derived from {achieved goals} that are substituted for the real goals of the task; this increases the frequency of positive rewards. {Hindsight experience replay} is used with off-policy RL algorithms, such as DQN~\cite{mnih2015human} and DDPG~\cite{lillicrap2015continuous}, for experience replay and has several extensions~\cite{schaul2015prioritized,liu2018competitive}. The present experiments show that simply applying HER with SIL does not lead to an agent capable of performing tasks from the Fetch robot environment. In summary, self-imitation learning with on-policy algorithms for tasks that require continuous control, and for which rewards are sparse, remains unsolved.

In this chapter, {episodic self-imitation learning (ESIL) for goal-oriented problems that provide only sparse rewards is proposed and combined with a state-of-the-art on-policy RL algorithm:} proximal policy optimisation (PPO). In contrast to standard SIL, which samples past good transitions from the replay buffer for imitation learning, {the proposed} ESIL adopts the entire set of current episodes {(successful or not)}, and modifies them into ``expert'' trajectories based on HER. An extra trajectory selection module is also introduced to relieve the effects of sample correlation~\cite{lee2019sample} in updating the network. Figure~\ref{fig:esil} shows the difference between naive SIL+HER and ESIL. During training by SIL+HER, a batch of transitions is sampled from the replay buffer; these are modified into ``hindsight experiences'' and used directly in self-imitation learning. In contrast, ESIL utilises the entire currently collected episodes and converts them into hindsight episodes. The trajectory selection module removes undesired transitions in the hindsight episodes. Using tasks from the OpenAI Fetch environment~\cite{plappert2018multi}, this chapter demonstrates that the proposed ESIL approach effectively trains the agent which is required to solve continuous control problems, and shows that it achieves state-of-the-art results on several tasks.

\begin{figure}[t]
\centering
\includegraphics[width=0.95\textwidth]{figures/chapter3/buffer.png}
\caption{Illustration of the difference between self-imitation learning (SIL)+hindsight experience replay (HER) and episodic self-imitation learning (ESIL).}
%% PPO+SIL+HER samples a batch of transitions from the replay buffer and modify them into hindsight experiences and conduct self-imitation learning directly. PPO+ESIL utilizes entire current collected episodes and convert them into hindsight episodes. Another trajectory selection module is used to remove undesired transitions in each hindsight episodes.
\label{fig:esil}
\end{figure}

The primary contribution of this chapter is a novel episodic self-imitation learning (ESIL) algorithm that can solve continuous control problems {in environments providing only sparse rewards}; in doing so, it also empirically answers an open question posed by Plappert \textit{et al.}~\cite{plappert2018multi} -- ``\textit{How to combine HER with state-of-the-art on-policy RL algorithms like
PPO?}". {The proposed ESIL approach also provides} a more efficient way to perform exploration in goal-oriented settings than the standard self-imitation learning algorithm. 
% Finally, to our knowledge, this approach achieves the best results for four moderately complex robot control tasks in simulation. 
Finally, this approach outperforms other on-policy RL algorithms in all simulated robot manipulation tasks.

% Related work
\section{Related Work}
% exploration
Imitation learning (IL) can be divided into two main categories: behavioural cloning and inverse reinforcement learning~\cite{hussein2017imitation}. Behavioural cloning involves the learning of behaviours from demonstrations~\cite{bojarski2016end,xu2017end,torabi2018behavioral}. Other extensions have an expert in the loop, such as DAgger~\cite{ross2011reduction}, or use an adversarial paradigm for the behavioural cloning method~\cite{ho2016generative,wang2017robust}. The inverse reinforcement learning estimates a reward model from expert trajectories~\cite{ng2000algorithms,abbeel2004apprenticeship,ziebart2008maximum}. Learning from demonstrations is powerful for complex robot manipulation tasks~\cite{finn2016guided,zhang2018deep,pmlr-v78-finn17a,rajeswaran2017learning,fang2019survey}. Prior work has used demonstrations to accelerate  learning~\cite{rajeswaran2017learning,vevcerik2017leveraging,nair2018overcoming},  nevertheless demonstrations are often collected by an expert policy or human actions. In contrast to these approaches, {episodic self-imitation learning (ESIL) does not need demonstrations.}

% self-learning and self-imitation learning
Self-imitation learning (SIL)~\cite{oh2018self} can be used for exploiting past experiences for parametric policies. It~has a similar flavor to ~\cite{gangwani2018learning, pmlr-v97-wu19a}, in that the agent learns from imperfect demonstrations. During~training, past good experiences are stored in the replay buffer. When SIL starts, transitions are sampled from the replay buffer according to the advantage values. {In the work of~\cite{tang2020self}, generalised~SIL was proposed as an extension of SIL. It uses an $n$-bound $Q$-learning approach to generalise the original SIL technique, and shows robustness to a wide range of continuous control tasks. Generalised SIL can also be combined with both deterministic and stochastic RL algorithms. Guo \textit{et al.}~\cite{guo2019self} points out that using imitation learning with past successful experiences could lead to a sub-optimal policy. Instead of imitating past good trajectories, a trajectory-conditioned policy~\cite{guo2019self} is proposed to imitate trajectories in diverse directions, encouraging exploration in environments where exploration is otherwise difficult.} Unlike SIL, episodic self imitation learning (ESIL) applies HER to the current episodes to create ``imperfect'' demonstrations for imitation learning; this also requires introducing a trajectory-selection module to reject undesired samples from the hindsight experiences. In the work of~\cite{lee2019sample}, it was shown that the agent benefits from using whole episodes in updates, rather~than uniformly sampling in the sparse or delayed reward environments. The present experiments suggest that episodic self-imitation learning {achieves better performance in an agent that must learn to perform continuous control in environments delivering sparse rewards}.

% Hindsight
Recently, the technique known as hindsight learning has been developed. Hindsight experience replay (HER)~\cite{andrychowicz2017hindsight} is an algorithm that can overcome the exploration problems in multi-goal environments, delivering sparse rewards. Hindsight policy gradient (HPG)~\cite{rauber2018hindsight} introduces techniques that enable the learning of goal-oriented policies using hindsight experiences with on-policy RL algorithms. However, the current implementation of HPG has only been evaluated for agents that need to perform discrete actions, and one drawback of hindsight policy gradient estimators is the computational cost because of the goal-oriented sampling. Unlike HPG, episodic self-imitation learning (ESIL) combines episodic hindsight experiences with imitation learning, which aids learning at the beginning of training. Furthermore, ESIL can be applied to continuous control, making it more suitable for control problems that demand greater precision.

% Methodology
\section{Methodology}
\label{sec:method}
The proposed method combines PPO and episodic self-imitation learning to maximally use hindsight experiences for exploration to improve learning. {Recent advantages in episodic update~\cite{lee2019sample} and hindsight experiences~\cite{andrychowicz2017hindsight} are also leveraged} to guide exploration for on-policy RL.

\subsection{Episodic Self-Imitation Learning}
The present method aims to use episodic hindsight experiences to guide the exploration of the PPO algorithm. To this end, hindsight experiences are created from current episodes. For an episode $i$, let there be $T$ timesteps; after $T$ timesteps, a series of transitions $\tau_i =\{ (s_{t}, a_{t}, r_{t}, s_{t+1}, g, g^{ac}_{t+1})\}_{t=0:T-1}$ is collected. If at timestep $t=T-1$, in $s_{T}$, $g^{ac}_{T} \neq g$, it implies that in this episode, the agent failed to achieve the original goal. Simply, to create hindsight experiences, {the achieved goal $g^{ac}_{T}$ in the last state $s_{T}$ is selected and considered as the modified desired goal $g^{\prime}$, i.e., $g^\prime = g^{ac}_{T}$.} Next, a new reward $r_{t}^{\prime}$ is computed under the new goal $g^\prime$. Then, {a new ``imagined'' episode is achieved}, and a new series of transitions $\tau_i^{\prime} = \{(s_{t}, a_{t}, r_{t}^{\prime},s_{t+1}, g^\prime, g^{ac}_{t+1})\}_{t=0:T-1}$ is collected.
Then, an approach to self-imitation learning based on episodic hindsight experiences is proposed, which applies the policy updates to both hindsight and in-environment episodes. Proximal policy optimisation (PPO) is used as the base RL algorithm, which is a state-of-the-art on-policy RL algorithm. With current and corresponding hindsight experiences, a new objective function is introduced and defined as:
\begin{equation}
  \centering
  \mathcal{L} = \alpha\mathcal{L}_{PPO} + \beta\mathcal{L}_{ESIL},
  \label{eq:loss}
\end{equation}
where $\alpha$ is the weight coefficient of $\mathcal{L}_{PPO}$. In the experiments, we set $\alpha=1$ as default {to balance the contribution of $\mathcal{L}_{PPO}$ and $\mathcal{L}_{ESIL}$}. $\mathcal{L}_{PPO}$ is the loss of PPO which can be written as:
\begin{equation}
  \mathcal{L}_{PPO} = \mathcal{L}_{policy} - c\mathcal{L}_{value}
  \label{eq:ppo_loss}
\end{equation}
where $\mathcal{L}_{policy}$ is the policy loss which is parameterised by $\theta$, $\mathcal{L}_{value}$ is the value loss which is parameterised by $\psi$, and $c$ is the weight coefficient of the $\mathcal{L}_{value}$, which is set to 1 to match the default PPO setting~\cite{schulman2017proximal}. The policy loss, $\mathcal{L}_{policy}$, can be represented as:
\begin{equation}
  \centering
  \mathcal{L}_{policy} = \mathbb{E}_{s_{t}, a_{t}, g\in\Tau}\biggr[\min \left (\frac{\pi(a_{t}|s_{t}, g;\theta)}{\pi(a_{t}|s_{t}, g;\theta_{old})}A_{t}, \clip \left (\frac{\pi(a_{t}|s_{t}, g;\theta)}{\pi(a_{t}|s_{t}, g;\theta_{old})}, 1-\epsilon, 1+\epsilon \right )A_{t}\right )\biggr],
  \label{eq:policy_loss}
\end{equation}
here, $A_{t}$ is the advantage value, and can be computed as $G_{t} - V(s_{t}, g;\psi)$. $V(s_{t}, g;\psi)$ is the state value at timestep $t$ which is predicted by the critic network. $G_{t}$ is the return at timestep $t$. $\epsilon$ is the clip ratio. $\Tau$~indicates original trajectories. {The value loss is a squared error loss $\mathcal{L}_{value}=\mathbb{E}[(V(s_{t}, g;\psi) - G_{t})^{2}]$.}

For the $\mathcal{L}_{ESIL}$ term, $\beta$ is an adaptive weight coefficient of $\mathcal{L}_{ESIL}$; it can be defined as the ratio of samples which are selected for self-imitation learning:
\begin{equation}
  \beta=\frac{N_{ESIL}}{N_{Total}},
\end{equation}
where $N_{ESIL}$ is the number of samples used for self-imitation learning and $N_{Total}$ is the total number of collected samples.
The episodic self-imitation learning loss $\mathcal{L}_{ESIL}$ can be written as: 
\begin{equation}
  \centering
  \mathcal{L}_{ESIL}=\mathbb{E}_{s_{t}, a_{t}, g^\prime\in\Tau^{'}}\left[\log\pi\left(a_{t}|s_{t}, g^\prime;\theta\right)\mathcal{F}_t\right],
  \label{eq:imitaion}
\end{equation}
where $\Tau^{'}$ denotes hindsight trajectories and $\mathcal{F}_t$ is the trajectory selection module which is based on returns of the current episodes, $G$, and the returns of corresponding hindsight experiences, $G^{\prime}$.

\subsection{Episodic Update with Hindsight}
Two crucial issues of ESIL are: 1) hindsight experiences are sub-optimal, and 2) the {detrimental} effect of updating networks with correlated trajectories. 
Although episodic self-imitation learning makes exploration more effective, hindsight experiences are not from experts and not ``perfect'' demonstrations.  With the training process continuing, if the agent is always learning these imperfect demonstrations, the policy will be stuck at the sub-optimal, or experience overfitting.

To prevent the agent from learning imperfect hindsight experiences, {hindsight experiences are actively selected based on returns.} With the same action, different goals may lead to different results. The proposed method only selects hindsight experiences that can achieve higher returns. The illustration of the trajectory selection module is in Figure~\ref{fig:hs_princple}. For an episodic experience and its hindsight experience, the returns of the episodic experience and its hindsight experience can be calculated, respectively. 
In a trajectory, at each timestep $t$, the return $G_{t}$ can be calculated by $G_{t} = G(s_t, g) = r_{t} + \gamma r_{t+1} + \gamma^{2} r_{t+2} + ... + \gamma^{T-t-1} r_{T-1}$. Then, for a trajectory $\tau_{i}$, we have $\left\{G_{0}^{i}, G_{1}^{i}, G_{2}^{i}, ..., G_{T-1}^{i}\right\}$.

For the hindsight experiences, similarly, the return $G_{t}'$ for each timestep $t$,~with respect to the hindsight goal $g^\prime$, can be calculated. Based on the modified trajectory $\tau_{i}^{\prime}$ with the same length of $\tau_{i}$, we therefore have the returns $\left\{G_{0}^{i^\prime}, G_{1}^{i^\prime}, G_{2}^{i^\prime}, ..., G_{T-1}^{i^\prime}\right\}$. During training, the~hindsight experiences with higher returns are used for self-imitation learning. The rest of the hindsight experiences will be \td{assumed to be of lower importance} and ignored.
Then, Equation~\eqref{eq:imitaion} can be rewritten as:
\begin{equation}
  \centering
  \mathcal{L}_{ESIL}=\mathbb{E}_{s_{t}, a_{t}, g^\prime\in\Tau^{'},g\in\Tau}\biggr[\log\pi_{\theta}\left(a_{t}|s_{t}, g^\prime\right)\mathcal{F}\left(s_{t}, g, g^\prime\right)\biggr],
  \label{eq:final_loss}
\end{equation}
where $\mathcal{F}\left(s_{t}, g, g^{\prime}\right)$ is the trajectory selection module. The selection function can be expressed as
\begin{equation}
  \centering
  \mathcal{F}\left(s_{t}, g, g^\prime \right)=\mathds{1}\left[G(s_{t}, g^\prime) > G(s_{t}, g)\right],
\end{equation}
here, {$\mathds{1}(\cdot)$ is the unit step function}. Consider the OpenAI Fetch environment as an example. For~a failed trajectory, the rewards $r_{t}$ are $\{-1,-1,\cdots, -1\}$. {The desired goal is modified} to construct a new hindsight trajectory, and the new rewards $r_{t}^{\prime}$ become $\{-1,-1,\cdots, 0\}$. Then,  {$G$ and $G^{\prime}$ can be calculated~separately.}
\begin{figure}[t!]
  \centering
  \includegraphics[width=\columnwidth]{figures/chapter3/hgs.png}
  \caption{A simplified illustration of trajectory selection. Blue trajectories indicate original experiences. Orange trajectories indicate hindsight experiences. Solid trajectories in the hindsight experiences are selected by the trajectory selection module of ESIL with new ``imagined'' goals.}
  \label{fig:hs_princple}
\end{figure}

From a goal perspective, episodic self-imitation learning (ESIL) tries to explore (desired) goals to get positive returns.~It can be viewed as a form of multi-task learning, {because ESIL has two objective functions to be optimised jointly.}~It is also related to self-imitation learning (SIL)~\cite{oh2018self}.~However,~ the~difference is that SIL uses $\left(G-V(s;\theta)\right)_{+}$ on past experiences to learn to choose the action chosen in the past in a given state, rather than goals. The full description of ESIL can be found in Algorithm~\ref{alg}.
% the algorithm
\begin{algorithm}[t!]
  \caption{Proximal Policy Optimisation (PPO) with Episodic Self-Imitation Learning (ESIL)}
  \label{alg}
  \begin{algorithmic}[1]
    \REQUIRE an actor network $\pi(s, g;\theta)$, a critic network $V(s, g;\psi)$, the maximum steps $T$ of an episode, a reward function $\mathcal{R}$, number of iterations $N$, number of episodes for each iteration $M$, weight coefficient of critic loss term $c$, weight coefficient of PPO term $\alpha$, adaptive coefficient $\beta$  
    %\STATE Initialize the parameters $\theta$ and $\psi$ of both actor and critic networks
    \FOR{$\mathrm{iteration} = 1, 2, \cdots N$}
    \STATE $\Tau \leftarrow \varnothing$, $\Tau^{\prime} \leftarrow \varnothing$ 
    \FOR{$\mathrm{episode} = 1, 2, ..., M$}
    \STATE $\tau \leftarrow \varnothing$
    \FOR{$t = 0, 1, \cdots, T-1$}
    \STATE Sample an action $a_{t}$ using the actor network $\pi(s_{t}, g;\theta)$
    \STATE Execute the action $a_{t}$ and observe a new state $s_{t+1}$
    \STATE Store the transition $\left(s_{t}, a_{t}, r_{t}, s_{t+1}, g^{ac}_{t+1}\right)$ in $\tau$
    \ENDFOR

    \FOR{\textbf{each} transition $\left(s_{t}, a_{t}, r_{t}, g, g^{ac}_{t+1} \right)$ in $\tau$}
    \STATE Clone the transition and replace $g$ with $g^{\prime}$, where $g^{\prime}=g^{ac}_{T}$
    \STATE $r_{t}^{\prime}:=\mathcal{R}\left(s_{t}, a_{t}, g^{\prime}\right)$
    \STATE Store the transition $\left(s_{t}, a_{t}, r_{t}^{\prime}, g^{\prime}, g^{ac}_{t+1}\right)$ in $\tau^{\prime}$
    \ENDFOR
    % next loop
    \STATE Store the trajectory $\tau$ and the hindsight trajectory $\tau^{\prime}$ in $\Tau$ and $\Tau^{\prime}$, respectively 
    \ENDFOR

    \STATE Calculate the Returns $G$ and $G^{\prime}$ for all transitions in $\Tau$ and $\Tau^{\prime}$, respectively
    \STATE Calculate the PPO loss: $\mathcal{L}_{PPO} = \mathcal{L}_{policy} - c\mathcal{L}_{value}$
    \STATE Calculate the ESIL loss: $\mathcal{L}_{ESIL}$ using Eq.\eqref{eq:imitaion}
    \STATE Update the parameters $\theta$ and $\psi$ using loss $\mathcal{L}=\alpha\mathcal{L}_{PPO} + \beta\mathcal{L}_{ESIL}$
    \ENDFOR 
\end{algorithmic}
\end{algorithm}

\section{Experiments}
\label{sec:experiment}
The proposed method is evaluated on several multi-goal environments, including the Empty Room environment and the OpenAI Fetch environment (see Figure~\ref{fig:env}). The Empty Room environment is a toy example, and has discrete action spaces. In the Fetch environment, there are four robot tasks with continuous action spaces. To obtain a comprehensive comparison between the proposed method and other baselines, suitable baselines are selected for different environments. In addition, ablation studies of the trajectory selection module are also performed.
\begin{figure}[t]
\minipage{0.19\textwidth}
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/empty_room.png}
  ({a}) Empty Room         
\endminipage\hfill
\minipage{0.19\textwidth}
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/reach.png}
  ({b}) Reach 
\endminipage\hfill
\minipage{0.19\textwidth}%
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/push.png}
  ({c}) Push      
\endminipage\hfill
\minipage{0.19\textwidth}%
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/pick.png}
  ({d}) Pick$\&$Place
\endminipage\hfill
\minipage{0.19\textwidth}%
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/slide.png}
  ({e}) Slide     
\endminipage
\caption{Evaluation environments. ({a}) is the Empty Room environment, in which a yellow circle indicates the position of the agent and a red star represents a target position. ({b}--{e}) are the tasks in the Fetch robot environment. The red spot represents a target position.}
\label{fig:env}
\end{figure}

\subsection{Environments}
\textbf{{Empty Room (grid-world) environment}}:
The Empty Room environment is a simple grid-world environment. The agent is placed in an $11\times11$ grid, representing the room. The goal of the agent is to reach a target position in the room. The start position of the agent is at the left upper corner of the room, and the target position is randomly selected  within the room. When the agent chooses an action that would fall outside the grid area, the agent will stay at the current position. The length of each episode is 32. 
The desired goal, $g$, is a two-dimensional grid coordinate, which represents the target position. The achieved goal, $g^{ac}_{t}$, is also a two-dimensional coordinate, which represents the current position of the agent, and finally, the observation is a two-dimensional coordinate which represents the current position of the agent. 
The agent has five actions: left, right, up, down, and stay; the agent executes a random action with a probability of 0.2. The agent can get $+1$ as a reward only when $g^{ac}_{t+1}=g$, otherwise, it gets a reward of $0$.

The agent is trained with 1 CPU core. In each epoch, 100 episodes are collected for the training. After each epoch, the agent is evaluated for 10 episodes. During training, the actions are sampled from the categorical distribution. During evaluation, the action with the highest probability will be chosen.

\textbf{{Fetch robot (continuous) environment}%.
}~\cite{plappert2018multi}:~The Fetch robot environment is a physically plausible simulation based on the real Fetch robot. This environment aims to provide a platform to tackle problems that are close to practical challenging robot manipulation tasks. Fetch is a 7-DoF robot arm with a two-finger gripper. The Fetch environment include four~tasks: Reach, Push, Pick$\&$Place and Slide. For all Fetch tasks, the length of each episode is 50. The~desired goal, $g$, is a three-dimensional coordinate that represents the target position. If~a task has an object, the achieved goal $g^{ac}_{t}$ is a three-dimensional coordinate that represents the position of the object. Otherwise, $g^{ac}_{t}$ is a three-dimensional coordinate that represents the position of the gripper.
Observations include the following information: position, velocity, and state of the gripper. If a task has an object, the position, velocity, and rotation information of the object is included. Therefore, the~observation of the Reach task is a 10-dimensional vector.~The observation of other tasks is a 25-dimensional vector.~The~action is a four-dimensional vector. The first three dimensions represent the relative position the gripper needs to move in the next step. The last dimension indicates the distance between the fingers of the gripper.
The reward function can be written as $r_{t}=-\mathds{1}\left(\left\|g^{ac}_{t+1} - g\right\|>\epsilon\right)$, where $\epsilon =0.05$.

In the Fetch environment, for the Reach, Push, and Pick$\&$Place tasks, the agent is trained using 16 CPU cores. In each epoch, 50 episodes are collected for training. The Slide task is more complex than others, so 32 CPU cores are used. In each epoch, 100 episodes are collected for training. The~{message passing interface (MPI) framework is used to perform} synchronization when updating the network. After each epoch, the agent is evaluated for 10 episodes by each MPI worker. Finally,~the~success rate of each MPI worker is averaged. During training, the actions are sampled from {multivariate normal} distributions. In the evaluation phase, the mean value of the Gaussian distribution is used as an action.

The proposed method, termed PPO+ESIL, is compared with different baselines in the different environments. All results are plotted based on five runs with different seeds. The solid line is the median value. The upper bound is the {75th} percentile, and the lower bound is the {25th} percentile.

\subsection{Experimental Settings}
\textbf{Network structure}: Both the actor network and the critic network have three hidden layers with 256~neurons. ReLU is selected as the activation function for the hidden layers. In the grid-world environment, the actor network builds a categorical distribution. In the Fetch environment, the actor network builds Gaussian {distribution by producing mean values and the standard deviations}.

\textbf{Hyperparameters}: For all experiments, the learning rate is 0.0003 for both the actor and critic networks. The discount factor $\gamma$ is 0.98. Adam is chosen as an optimiser. For each epoch, the~actor network and critic network are updated 10 times. The clip ratio of the PPO algorithm is 0.2. For the grid-world environment, networks are trained for 100 epochs with a batch size equal to 160. Each~epoch consists of 100 episodes. For the Fetch environment, in the Reach task, networks are trained for 100 epochs and other tasks for 1000 epochs with a batch size equals to 125. For the Reach, Push, and Pick$\&$Place tasks, each epoch consists of 50 episodes. For the Slide task, each epoch consists of 100 episodes. In designing the experiments, the number of episodes within an epoch is a balance between being able to train, the length of time required to run experiments and the maximum number of timesteps that would be required to achieve a goal. All environments have a fixed maximum number of timesteps, but this maximum differs depending on the problem or environment. This means that the number of state--action pairs can differ between two environments that have the same number of episodes and the same number of epochs. We arrange the episodes to try to compensate for the number of state--action pairs collected during training to make experiments easier to compare. The models are trained on a machine with an Intel i7-5960X CPU and 64GB RAM.

\begin{figure}[h!]
\minipage{0.5\textwidth}
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/empty_room_baseline.pdf}
  ({a}) Comparison with on-policy baselines\hspace{3em} 
\endminipage
\minipage{0.5\textwidth}
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/empty_room_hs.pdf}
  ({b}) {Ablation study of selection module}
\endminipage\hfill
\minipage{0.5\textwidth}%
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/empty_room_samples.pdf}
  ({c}) Variation of adaptive weight $\beta$
\endminipage
\minipage{0.5\textwidth}%
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/empty_room_her.pdf}
  ({d}) Comparison with off-policy baselines
\endminipage\hfill
\caption{Results of the grid-world environment. ({a}) Comparing the performance of PPO+ESIL between the on-policy approaches. ({b}) An ablation study on the trajectory selection module. ({c}) The variation of adaptive weight coefficient $\beta$ through training. ({d}) Comparison of the performance of PPO+ESIL to an off-policy approach: DQN+HER.}
\label{fig:toy_results}
\end{figure}

%\section{Results and Discussions}
\subsection{Results on Grid-World Environment}
To understand the basic properties of the proposed method, the toy Empty Room environment is used to evaluate ESIL.
%We compare the proposed method, named as PPO+HSL, with other baselines on the grid-world environment.
The following baselines are considered:
\begin{itemize}
    \item PPO: vanilla PPO~\cite{schulman2017proximal} for discrete action spaces;
    \item PPO+SIL/PPO+SIL+HER: Self-imitation learning (SIL) is used with PPO to solve hard exploration environments by imitating past good experiences~\cite{oh2018self}. In order to solve sparse rewards tasks, hindsight experience replay (HER) is applied to sampled transitions;
    \item DQN+HER: Hindsight experience replay (HER), designed for sparse reward problems, is~combined with a deep Q-learning network (DQN)~\cite{andrychowicz2017hindsight}; this is an off-policy algorithm;
    \item Hindsight Policy Gradients (HPG): the vanilla implementation of HPG that is only suitable for discrete action spaces~\cite{rauber2018hindsight}.
\end{itemize}
More specifically, {PPO+ESIL is compared with the above baseline methods} in Figure~\ref{fig:toy_results}a. This shows that PPO+ESIL converges faster than the other four baselines, and PPO+SIL converges faster than vanilla PPO, because PPO+SIL reuses past good experiences to help exploration and training. Hindsight policy gradient (HPG) is slower than the others because goal sampling is not efficient and is unstable.

The performance of the trajectory selection module is evaluated in Figure~\ref{fig:toy_results}b. It shows that the selection strategy helps improve the performance. Hindsight experiences are not always perfect; the trajectory selection module filters some undesirable, modified experiences. Through adopting this selection strategy, the chance of agents learning from poor trajectories is reduced. The adaptive weight coefficient, $\beta$, is also investigated in these experiments. {In Figure~\ref{fig:toy_results}c, it can be seen that at the initial stages of training, $\beta$ is high}. {This is because, at this stage, the agent very seldom achieves the original goals}. The hindsight experiences can yield higher returns than the original experiences. {Therefore,~a~large proportion of hindsight experiences is selected to conduct self-imitation learning, helping the agent learn a policy for moving through the room}. In the later stages of training, the agent can achieve success frequently, and the hindsight experiences might be redundant (e.g., $G(s_{t}, g)\geq G(s_{t}, g^{\prime})$). In~this case, undesired hindsight experiences are removed by using the trajectory selection module and $\mathcal{L}_{PPO}$ leads the training. However, when the trajectory selection module is not employed, all hindsight experiences are used throughout the training, including the redundant hindsight experiences. This leads to overfitting and makes training unstable. Thus, $\mathcal{L}_{ESIL}$ can provide the agent with a better initial policy, and the adaptive weight coefficient $\beta$ can balance the contributions of $\mathcal{L}_{PPO}$ and $\mathcal{L}_{ESIL}$ properly during training.

Finally, {the combination of PPO+ESIL is also compared with DQN+HER,} which is an off-policy RL algorithm, in Figure~\ref{fig:toy_results}d. It shows that DQN+HER works a little better than ESIL at the {start of training}. However, the proposed method achieves similar results to DQN+HER later in training.

\subsection{Results on Continuous Control Environment}
% continuous control
{Continuous control problems are generally more challenging for reinforcement learning. In the experiments of this section, the aim is to investigate how useful the proposed method is for several hard exploration Fetch tasks. These robot manipulation tasks are commonly used to assess the performance of RL methods for continuous control.} The following baselines are considered: 
\begin{itemize}
    \item PPO: the vanilla PPO~\cite{schulman2017proximal} for continuous action spaces;
    \item PPO+SIL/PPO+SIL+HER: Self-imitation learning is used with PPO to solve hard exploration environments by imitating past good experiences~\cite{oh2018self}. For sparse rewards tasks, hindsight experience replay (HER) is applied to sampled transitions;
    \item DDPG+HER: this is the state-of-the-art off-policy RL algorithm for the Fetch tasks. Deep~deterministic policy gradient (DDPG) is trained with HER to deal with the sparse reward problem~\cite{andrychowicz2017hindsight}.
\end{itemize}

\begin{figure}[h!]
\centering
\minipage{0.5\textwidth}
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/reach_baseline.pdf}
  ({a}) Reach
\endminipage
\minipage{0.5\textwidth}%
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/push_baseline.pdf}
  ({b}) Push
\endminipage\hfill
\minipage{0.5\textwidth}%
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/pick_baseline.pdf}
  ({c}) Pick$\&$Place
\endminipage
\minipage{0.5\textwidth}%
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/slide_baseline.pdf}
  ({d}) Slide
\endminipage\hfill
\caption{Results of comparison between ESIL and on-policy baselines in all Fetch tasks. \td{From the results, ESIL is able to resolve all of the tasks, while other baselines can only solve the simple Reach task.}}
\label{fig:baseline_compare}
\end{figure}

\subsubsection{Comparison to On-Policy Baselines}
{Figure~\ref{fig:baseline_compare}, PPO+ESIL} achieves reasonable results in all Fetch tasks. {In contrast}, PPO, PPO+SIL, and PPO+SIL+HER do not work in all tasks, except for the Reach task. In comparison with the other selected tasks from the Fetch environment, Reach task is relatively simple, because there is no object to be manipulated. For other tasks, it is quite difficult for the agent to achieve sufficient positive rewards during exploration, because of their rare occurrence. Although PPO+SIL utilises past good experiences to help exploration, it is still faced with the difficulty that past experiences do not easily achieve positive rewards. From the experiments ({see Figure~\ref{fig:baseline_compare}}), PPO+SIL (no hindsight) converges much more slowly than using PPO only. Attempting to use only the original trajectories for self-imitation learning leads to unsatisfactory performance. For PPO+SIL+HER (with no episodic update), the sampled transitions are modified into hindsight experiences, achieving better performance in the Reach and Slide tasks. However, this transition-based method still cannot solve the other two manipulation tasks. In contrast, the proposed PPO+ESIL, utilising episodic hindsight experiences from failed trajectories, can quickly achieve positive rewards at the start of~training.

\begin{figure}[h!]
\centering
\minipage{0.5\textwidth}
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/reach_hs.pdf}
  ({a}) Reach
\endminipage
\minipage{0.5\textwidth}%
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/push_hs.pdf}
  ({b}) Push
\endminipage\hfill
\minipage{0.5\textwidth}%
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/pick_hs.pdf}
  ({c}) Pick$\&$Place
\endminipage
\minipage{0.5\textwidth}%
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/slide_hs.pdf}
  ({d}) Slide
\endminipage\hfill
\caption{Results of ablation studies with and without using the trajectory selection module in all Fetch tasks. \td{In three challenging tasks: Push, Pick$\&$Place and Slide, the trajectory selection module can stabilise the training and prevent performance degradation in two of three cases (Push and Pick$\&$Place).}}
\label{fig:hs_compare}
\end{figure}

\begin{figure}[h!]
\centering
\minipage{0.5\textwidth}
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/reach_samples.pdf}
  ({a}) Reach
\endminipage
\minipage{0.5\textwidth}%
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/push_samples.pdf}
  ({b}) Push
\endminipage\hfill
\minipage{0.5\textwidth}%
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/pick_samples.pdf}
  ({c}) Pick$\&$Place
\endminipage
\minipage{0.5\textwidth}%
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/slide_samples.pdf}
  ({d}) Slide
\endminipage\hfill
\caption{Variation in adaptive weight coefficient $\beta$ through the training in all Fetch tasks. \td{In the early stage of training, the weight of the ESIL term is higher, which indicates more samples need to be modified to facilitate the exploration. As training progresses, the weight decreases, which means that the agent can achieve more positive samples and less modified samples are required.}}
\label{fig:her_samples_compare}
\end{figure}

\subsubsection{Ablation Study of Trajectory Selection Module}
To investigate the effect of trajectory selection, {ablation studies are performed} to validate the selection strategy of our approach. {Figure~\ref{fig:hs_compare}}, when the trajectory selection module is not used, the~ performance of the agent increases at first, and then starts to decrease. This suggests that the agent starts to converge to a sub-optimal location. However, {Figure~\ref{fig:hs_compare}d}, for the Slide task, the agent converges faster without the trajectory selection module and has better performance. {This is likely to be because the Slide task is the most difficult of the Fetch environment. During training, the agent is unlikely to achieve positive rewards. Figure~\ref{fig:her_samples_compare} also indicates that the value of $\beta$ in the Slide task is higher than values in other tasks, which means the majority of hindsight experiences have higher returns than original experiences. Thus, using {\em more} hindsight experiences (without filtering) accelerates training at this stage.} Nonetheless, the trajectory selection module prevents the agent from overfitting the hindsight experience in the other three tasks. Figure~\ref{fig:her_samples_compare}, shows the adaptive weight coefficient $\beta$ in all Fetch tasks. When the trajectory selection module is used, the value of $\beta$ decreases with the increase in training epochs. This implies that the agent can achieve a greater proportion of the original goals in the latter stages of training, and fewer hindsight experiences are required for self-imitation learning.
\begin{figure}[h!]
\minipage{0.5\textwidth}
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/reach_her.pdf}
  ({a}) Reach
\endminipage
\minipage{0.5\textwidth}%
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/push_her.pdf}
  ({b}) Push
\endminipage\hfill
\minipage{0.5\textwidth}%
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/pick_her.pdf}
  ({c}) Pick$\&$Place
\endminipage
\minipage{0.5\textwidth}%
  \centering
  \includegraphics[width=\linewidth]{figures/chapter3/slide_her.pdf}
  ({d}) Slide
\endminipage\hfill
\caption{Results of the comparison between PPO+ESIL and DDPG+HER in all fetch tasks. \td{From the results, PPO+ESIL has worse sample efficiency than the off-policy baseline. However, PPO+ESIL achieves similar or even better performance in the end of training.}}
\label{fig:her_compare}
\end{figure}

\subsubsection{Comparison to Off-Policy Baselines}
Finally, {the proposed method is also compared} with a state-of-the-art off-policy algorithm: DDPG+HER. From {Figure~\ref{fig:her_compare}, it may be seen that} DDPG+HER converges faster than PPO+ESIL in all tasks. {However, PPO+ESIL obtains a similar performance to DDPG+HER. This is because DDPG+HER is an off-policy algorithm and uses a large number of hindsight experiences. A replay buffer is also employed to store  samples collected in the past. This approach has better sample efficiency than on-policy algorithms such as PPO.} Even so, {Figure~\ref{fig:her_compare}c} shows that PPO+ESIL still outperforms DDPG+HER in the Pick$\&$Place task and the success rate is close to 1. {This suggests that PPO+ESIL approximates the characteristics of on-policy algorithms, which have low sample efficiency, but are able to obtain a comparable performance to off-policy algorithms in continuous control tasks~\cite{schulman2017proximal}.}

\subsection{Overall Performance}
{Table~\ref{tb:ch3_last}} shows the average success rate of the last 10 epochs during training of baseline methods and PPO+ESIL. The proposed ESIL achieves the best performance in four out of five tasks. However,~PPO and PPO+SIL only obtain reasonable results in the Empty Room and Reach task. With the assistance of HER, PPO+SIL+HER obtains a better performance in the Slide task. {For the off-policy method of DDPG+HER, all five tasks are achieved, but it only obtain a better performance than PPO+ESIL in the Push task}.

\begin{table}[h]
  \centering
  \resizebox{\textwidth}{!}{
  \begin{tabular}{llllll}
    \toprule
           & {Empty Room} & {Reach} & {Push} & {Pick$\&$Place} & {Slide}  \\
    \midrule
    PPO & 1.000 $\pm$ 0.000 & 1.000 $\pm$ 0.000& 0.070 $\pm$ 0.001 & 0.033 $\pm$ 0.001 & 0.077 $\pm$ 0.001 \\
    PPO + SIL & 0.998 $\pm$ 0.002 & 0.225 $\pm$ 0.016 & 0.071 $\pm$ 0.001 & 0.036 $\pm$ 0.002 & 0.011 $\pm$ 0.001 \\
    PPO + SIL + HER &0.996 $\pm$ 0.013 & 1.000 $\pm$ 0.000 & 0.066 $\pm$ 0.011 & 0.035 $\pm$ 0.004 & 0.276 $\pm$ 0.011 \\
    DQN + HER &1.000 $\pm$ 0.000 & - & - & - & - \\
    DDPG + HER & - & 1.000 $\pm$ 0.000 & \textbf{0.996 $\pm$ 0.001} & 0.888 $\pm$ 0.008 & 0.733 $\pm$ 0.013\\
    HPG &0.964 $\pm$ 0.012 & - & - & - & - \\
    \midrule
    PPO + ESIL (Ours) & \textbf{1.000 $\pm$ 0.000} & \textbf{1.000 $\pm$ 0.000}& 0.984 $\pm$ 0.003& \textbf{0.986 $\pm$ 0.002}& \textbf{0.812 $\pm$ 0.015}\\
    \bottomrule
  \end{tabular}}
   \caption{Average success rate $\pm$ standard error in the last 10 epochs over five random seeds in all~environments {(\textbf{bold} indicates the best result among all methods).}}
  \label{tb:ch3_last}
\end{table}
% Conclusion

\section{Discussion}
\label{sec:conclusion}
This chapter proposed a novel method for self-imitation learning (SIL), in which an on-policy RL algorithm uses episodic modified trajectories, i.e., hindsight experiences, to update policies. Compared with standard self-imitation learning, episodic self-imitation learning (ESIL) has a better performance in continuous control tasks where rewards are sparse. As far as we know, it is also the first time that hindsight experiences have been combined with state-of-the-art on-policy RL algorithms, such as PPO, to solve relatively hard exploration tasks with continuous action spaces.

The experiments that we have conducted suggest that simply using self-imitation learning with the PPO algorithm, even with hindsight experience, leads to disappointing performance in continuous control Fetch tasks. In contrast, the episodic approach we take with ESIL is able to learn in these sparse reward settings. The auxiliary trajectory selection module and the adaptive weight $\beta$ help the training process to remove undesired experiences and balance the contributions to learning between the PPO term and the ESIL term automatically, and also increase the stability of training.

Our experiments suggest that the selection module is useful to prevent overfitting to sub-optimal hindsight experiences, but also that it does not always lead to learning a better policy faster. Despite~this, selection filtering appears to support learning a useful policy in challenging environments. The experiments we have conducted to date have utilised relatively small networks, and it would be appropriate to extend the experiments to consider more complex observation spaces, and to actor/critic networks, which are consequently more elaborate.

% Future work includes extending {the proposed} method to support hierarchical reinforcement learning (HRL) algorithms for more complex manipulation control tasks, such as in-hand manipulation. Episodic self-imitation learning (ESIL) can also be applied to learn sub-goal policies simultaneously.