### 2022-04-23 dedication/copyright.tex

\cleardoublepage

\begin{mycopyright}
The copyright of this thesis rests with the author. Unless otherwise indicated,
its contents are licensed under a Creative Commons Attribution-Non
Commercial 4.0 International Licence (CC BY-NC).
Under this licence, you may copy and redistribute the material in any medium
or format. You may also create and distribute modified versions of the work.
This is on the condition that: you credit the author and do not use it, or any
derivative works, for a commercial purpose.
When reusing or sharing this work, ensure you make the licence terms clear to
others by naming the licence and linking to the licence text. Where a work has
been adapted, you should indicate that the work has been changed and
describe those changes.
Please seek permission from the copyright holder for uses of this work that are
not included in this licence or permitted under UK Copyright Law.

\hfill Tianhong Dai
\end{mycopyright}