### 2022-08-21 acknowledgements/acknowledgements.tex

\cleardoublepage

\addcontentsline{toc}{chapter}{Acknowledgements}

\begin{acknowledgements}

I would like to express my deep gratitude to:

\begin{itemize}
 \item Anil Anthony Bharath: Anil is my supervisor and brings me into an academic research career. During my Ph.D. study, he provided me with plentiful guidance, help and trust in my research. In addition, he has also provided me with valuable suggestions on my career planning. Without him, I would not have been able to achieve the current results.
 \vspace*{3mm}
 \item Kai Arulkumaran: Kai is my closest colleague/collaborator in the BICI group. He has been a constant and continuing collaborator. In addition, he also has provided me many suggestions in daily life. He is also my ``leader" in the field of reinforcement learning.
 \vspace*{3mm}
 \item BICI Group Members: I would like to express my gratitude for their daily support.
 \vspace*{3mm}
 \item Meng Fang: Dr. Meng Fang was my mentor during my internship at Tencent AI Lab and he helped me a lot in the research of deep reinforcement learning.
 \item Shanxin Yuan: Dr. Shanxin Yuan was my mentor during my internship at Huawei Noah's Ark Lab and he helped me a lot in the research of low-level computer vision.
 \vspace*{3mm}
 \item My Parents: From the day I was born, my parents have provided me with unfailing love and support. Without them, I wouldn't be here today.
 \vspace*{3mm}
 \item My Wife: From the day we meet, she continually inspires me to pursue my dreams and gives me much love and care. During our two years together in the UK, she has been there to guide me through difficult times, to help me with my worries, and to allow me to focus on my work. She makes my life so full of hope and joy that no words can express my gratitude and love for her.
\end{itemize}

\end{acknowledgements}